---
title: "Standardization"
subtitle: "Understanding Confounding and Population Structure"
author: "Erik Westlund"
date: today
format:
  revealjs:
    theme: simple
    transition: slide
    slide-number: true
    chalkboard: false
    incremental: false
    code-line-numbers: false
    fig-align: center
    fig-width: 10
    fig-height: 6
    include-after-body: resources/title-logos.html
execute:
  echo: true
  warning: false
  message: false
---

```{r}
#| label: setup
#| include: false

library(framework)
scaffold()

# Source base population FIRST (creates base_population object)
source("scripts/00_base_population.R")

# Source all three simulations (each uses base_population)
source("scripts/sim1_age_only.R")
source("scripts/sim2_age_sex.R")
source("scripts/sim3_age_sex_goc.R")

# Set theme for plots
theme_set(theme_minimal(base_size = 14))
```

## Motivating Example: Delayed Seeking of Necessary Healthcare


::: {.incremental}
- **Motivation:** People wait to get help until a health issue becomes serious or avoidable harm has already occurred  
  - Skipping routine visits or screenings until symptoms worsen  
  - Delaying care due to cost, time, or caregiving demands  
  - Arriving at the health system only during a crisis  
- **Importance:** Reveals inequities in access, prevention, and confidence in the health system  
- **Question For Us:** Who faces greater risk (e.g., men or women?) and how can we compare that risk fairly?
:::

## The Challenge

Crude rates can mislead us about *who* is at risk and *why*


## Meeting Themes

Refresher:

::: {.incremental}
- **Sex-disaggregated data** tells us who (counts)
- **Gendered analysis** asks why — how roles, norms, and power shape health and behavior
- Those processes interact with age and the gender equality context
:::

::: {.fragment}
<br>

**This lecture's focus:** How **causal structure** and **population composition** determine what crude rates actually mean
:::

## Agenda

::: {.incremental}
1. **The core idea**: Statistical expectation and population structure
2. **Example 1**: Age → Outcome (composition matters)
3. **Example 2**: Age + Sex → Outcome (Simpson's paradox)
4. **Example 3**: Age + Sex + Gender Opportunity → Outcome (limits of manual standardization)
5. **Summary**: Getting the expected value right
:::

## The Core Problem

::: {.incremental}
- Every population has its own **mix of people** — by age, gender, opportunity, and access.  
- That mix shapes what we see in the data: who gets sick, who shows up, and when.  
- When we compare groups without accounting for those differences, we risk blaming behavior or biology for what is really structure.  
:::

---

## Statistical Expectation 

::: {.incremental}
- The **expected rate** is what we’d see if we looked across everyone in a group, *given their mix of ages and circumstances.*  
- If two groups have different mixes, their expected rates aren’t directly comparable.  
- We get the **wrong picture** because we’re averaging over different kinds of people.  
:::

---

## How Standardization Fixes It

::: {.incremental}
- **Step 1:** Agree on a common mix — the same age, gender, or opportunity structure for everyone.  
- **Step 2:** Recalculate each group’s rate *as if they had that same mix.*  
- **Result:** Differences now reflect real risk, not who happens to be older, poorer, or more constrained.  
:::

::: {.fragment}
<br>

**In short:** Standardization answers,  “What would these rates look like if all groups lived under the same conditions?”
:::

# Example 1: Age → Outcome {.section-break background-color="#2C3E50"}

## The Simplest Case: Age Only

**Research question:** What's the delayed healthcare seeking rate in our population?

::: {.incremental}
- We have simulated registry data on preventable health crises
- E.g., diabetic emergencies, hypertensive crises, or late-stage cancers
- **We define the risk by age:** Young: 5%, Middle: 15%, Older: 40%
- **Question:** What crude rate do we expect?
:::

::: {.fragment}
<br>

**Key insight:** Depends entirely on population age structure!
:::

## Causal Structure: Age → Outcome

```{r}
#| label: dag1
#| echo: false
#| fig-width: 6
#| fig-height: 3

library(ggdag)

dag1 <- dagify(
  Out ~ Age,
  coords = list(
    x = c(Age = 1, Out = 2),
    y = c(Age = 1, Out = 1)
  )
)

ggdag(dag1, text_size = 5, node_size = 20) +
  theme_dag() +
  labs(title = "Causal Structure: Age Determines Outcome")
```

::: {.fragment}
**No confounding here**. Just one cause: Age

But crude rate still depends on age.
:::

## Simulated Dataset 1: Age Distribution

```{r}
#| label: age-dist-sim1
#| echo: false

sim1_data %>%
  count(age_group) %>%
  mutate(prop = n / sum(n)) %>%
  ggplot(aes(x = age_group, y = prop, fill = age_group)) +
  geom_col(width = 0.6) +
  geom_text(aes(label = percent(prop, accuracy = 1)),
            vjust = -0.5, size = 6) +
  scale_y_continuous(labels = percent_format(), limits = c(0, 0.45)) +
  scale_fill_manual(values = c("#3498DB", "#E67E22", "#E74C3C")) +
  labs(
    title = "Population Age Structure (Dataset 1)",
    subtitle = "Known From Simulation P(Age) = (30%, 35%, 35%)",
    x = "Age Group",
    y = "Proportion"
  ) +
  theme(legend.position = "none")
```

## True Age-Specific Rates (Known)

```{r}
#| label: true-rates-sim1
#| echo: false

true_rates_sim1 <- data.frame(
  age_group = factor(c("Young", "Middle", "Older"),
                    levels = c("Young", "Middle", "Older")),
  true_rate = c(0.05, 0.15, 0.40)
)

ggplot(true_rates_sim1, aes(x = age_group, y = true_rate, fill = age_group)) +
  geom_col(width = 0.6) +
  geom_text(aes(label = percent(true_rate, accuracy = 1)),
            vjust = -0.5, size = 6) +
  scale_y_continuous(labels = percent_format(), limits = c(0, 0.45)) +
  scale_fill_manual(values = c("#3498DB", "#E67E22", "#E74C3C")) +
  labs(
    title = "True Age-Specific Rates (We set it)",
    subtitle = "These are the 'truth' we're trying to understand",
    x = "Age Group",
    y = "Late Presentation Rate"
  ) +
  theme(legend.position = "none")
```

## Computing the Expected Value

Using the expectation formula:

$$
E[Y] = \sum_{\text{age}} E[Y \mid \text{Age}] \times P(\text{Age})
$$

::: {.fragment}
```{r}
#| label: compute-expected-sim1
#| echo: true

# Our age-specific rates (known truth)
rates <- c(Young = 0.05, Middle = 0.15, Older = 0.40)

# Our population weights
weights <- c(Young = 0.30, Middle = 0.35, Older = 0.35)

# Expected value
expected_rate <- sum(rates * weights)
expected_rate
```
:::

::: {.fragment}
**Expected crude rate:** `r percent(expected_rate, accuracy = 0.1)`

This is what we should see in the data.
:::

## Observed vs Expected

```{r}
#| label: crude-vs-expected-sim1
#| echo: false

observed_crude <- mean(sim1_data$late_presentation)

comparison_sim1 <- data.frame(
  type = c("Expected\n(from formula)", "Observed\n(crude rate)"),
  rate = c(expected_rate, observed_crude)
)

ggplot(comparison_sim1, aes(x = type, y = rate, fill = type)) +
  geom_col(width = 0.6) +
  geom_text(aes(label = percent(rate, accuracy = 0.1)),
            vjust = -0.5, size = 7) +
  scale_y_continuous(labels = percent_format(), limits = c(0, 0.25)) +
  scale_fill_manual(values = c("#27AE60", "#3498DB")) +
  labs(
    title = "Expected vs Observed Crude Rate",
    subtitle = "Close match confirms: crude rate = weighted average by population structure",
    x = NULL,
    y = "Late Presentation Rate"
  ) +
  theme(legend.position = "none")
```

::: {.fragment}
**Key insight:** Crude rate reflects population age structure. Different population → different crude rate.
:::

## What If We Had a Different Population?

::: {.incremental}
- Imagine Population A: Mostly young people → low crude rate
- Imagine Population B: Mostly older people → high crude rate
- **Same age-specific rates** (5%, 15%, 40%)
- **Different crude rates** because $\text{Age}$ differs
:::

::: {.fragment}
<br>

**This is why standardization matters!**

To compare Population A vs B fairly, we need to apply their age-specific rates to a **standard population** and use the **same age weights** for both
:::

# Example 2: Age + Sex {.section-break background-color="#2C3E50"}

## Adding Gender: Simpson's Paradox

**Research question:** Do men or women have higher late-presentation rates?

::: {.incremental}
- Now we have Sex and Age in our data
- **Known truth** (from simulation):
  - Men have higher risk within **every** age group (by a few percentage points)
  - Women live longer → more older women
- What will crude rates show?
:::

## Causal Structure: Age + Sex → Outcome

```{r}
#| label: dag2
#| echo: false
#| fig-width: 8
#| fig-height: 4

dag2 <- dagify(
  Out ~ Age + Sex,
  Age ~ Sex,
  coords = list(
    x = c(Sex = 1, Age = 2, Out = 3),
    y = c(Sex = 2, Age = 1, Out = 2)
  )
)

ggdag(dag2, text_size = 5, node_size = 20) +
  theme_dag() +
  labs(title = "Causal Structure: Sex → Age → Outcome, Sex → Outcome")
```

::: {.fragment}
**This is confounding!**

Sex affects Age (women live longer), and Age affects Outcome.

Crude comparison of men vs women is **biased** by age composition.
:::

## Simulated Dataset 2: Age Structure by Sex

```{r}
#| label: age-by-sex-sim2
#| echo: false

sim2_data %>%
  count(sex, age_group) %>%
  group_by(sex) %>%
  mutate(prop = n / sum(n)) %>%
  ggplot(aes(x = age_group, y = prop, fill = sex)) +
  geom_col(position = "dodge") +
  geom_text(aes(label = percent(prop, accuracy = 1)),
            position = position_dodge(width = 0.9),
            vjust = -0.5, size = 4) +
  scale_y_continuous(labels = percent_format(), limits = c(0, 0.6)) +
  scale_fill_manual(values = c("#3498DB", "#E74C3C")) +
  labs(
    title = "Age Distribution by Sex (Dataset 2)",
    subtitle = "Women concentrated in older ages; men in younger ages",
    x = "Age Group",
    y = "Proportion",
    fill = "Sex"
  ) +
  theme(legend.position = "top")
```

::: {.fragment}
**Key:** P(Age | Man) ≠ P(Age | Woman), but **marginal P(Age) same as Dataset 1**
:::

## Crude Rates by Sex

```{r}
#| label: crude-sex-sim2
#| echo: false

crude_sex <- sim2_data %>%
  group_by(sex) %>%
  summarise(crude_rate = mean(late_presentation))

ggplot(crude_sex, aes(x = sex, y = crude_rate, fill = sex)) +
  geom_col(width = 0.6) +
  geom_text(aes(label = percent(crude_rate, accuracy = 0.1)),
            vjust = -0.5, size = 8) +
  scale_y_continuous(labels = percent_format(), limits = c(0, 0.30)) +
  scale_fill_manual(values = c("#3498DB", "#E74C3C")) +
  labs(
    title = "Crude Late-Presentation Rates by Sex",
    subtitle = "Women appear worse overall...",
    x = NULL,
    y = "Late Presentation Rate"
  ) +
  theme(legend.position = "none")
```

::: {.fragment}
**Observation:** Women have higher crude rate

**But is this the truth?**
:::

## Age-Specific Rates by Sex

```{r}
#| label: age-specific-sex-sim2
#| echo: true

age_sex_rates <- sim2_data %>%
  group_by(sex, age_group) %>%
  summarise(rate = mean(late_presentation), .groups = "drop")

age_sex_rates
```

::: {.fragment}
**Within every age group, men have higher rates!**

This is **Simpson's Paradox**.
:::

## Visualizing Simpson's Paradox

```{r}
#| label: paradox-viz-sim2
#| echo: false

# Ensure age_group is an ordered factor for line plotting
age_sex_rates_plot <- age_sex_rates %>%
  mutate(age_group = factor(age_group, levels = c("Young", "Middle", "Older"), ordered = TRUE))

ggplot(age_sex_rates_plot, aes(x = age_group, y = rate,
                          color = sex, group = sex)) +
  geom_line(linewidth = 1.5) +
  geom_point(size = 4) +
  geom_text(aes(label = percent(rate, accuracy = 0.1)),
            vjust = -1, size = 4) +
  scale_y_continuous(labels = percent_format(), limits = c(0, 0.50)) +
  scale_color_manual(values = c("#3498DB", "#E74C3C")) +
  labs(
    title = "Within Each Age Group, Men Have Higher Rates",
    subtitle = "Simpson's Paradox: Composition flips the overall pattern",
    x = "Age Group",
    y = "Late Presentation Rate",
    color = "Sex"
  ) +
  theme(legend.position = "top")
```

## What Went Wrong?

Remember the expectation formula:

$$
E[Y \mid \text{Sex}] = \sum_{\text{age}} E[Y \mid \text{Sex}, \text{Age}] \times P(\text{Age} \mid \text{Sex})
$$

::: {.fragment}

- $P(\text{Age} \mid \text{Man})$ is different from $P(\text{Age} \mid \text{Woman})$
- Women have more weight on **older ages** (where everyone has high rates)
- Men have more weight on **younger ages** (where everyone has low rates)
- Men have higher risk within every age group
- **Different age weights** flip the crude pattern despite consistent within-age differences
:::

## Indirect Standardization

**Goal:** What would the **expected number of events** be if each group had standard rates?

```{r}
#| label: indirect-std-sim2
#| echo: true

# Step 1: Use overall population rates as standard
standard_rates <- sim2_data %>%
  group_by(age_group) %>%
  summarise(std_rate = mean(late_presentation))

# Step 2: Calculate expected events for each sex using standard rates
indirect_std <- sim2_data %>%
  group_by(sex, age_group) %>%
  summarise(n = n(), .groups = "drop") %>%
  left_join(standard_rates, by = "age_group") %>%
  group_by(sex) %>%
  summarise(
    observed = sum(n * age_sex_rates$rate[match(paste(sex, age_group),
                                                  paste(age_sex_rates$sex, age_sex_rates$age_group))]),
    expected = sum(n * std_rate),
    SMR = observed / expected
  )

indirect_std
```

## SMR: Standardized Morbidity Ratio

```{r}
#| label: smr-viz-sim2
#| echo: false

ggplot(indirect_std, aes(x = sex, y = SMR, fill = sex)) +
  geom_col(width = 0.6) +
  geom_hline(yintercept = 1, linetype = "dashed", color = "black", linewidth = 1) +
  geom_text(aes(label = sprintf("%.2f", SMR)),
            vjust = -0.5, size = 8) +
  scale_y_continuous(limits = c(0, 1.5)) +
  scale_fill_manual(values = c("#3498DB", "#E74C3C")) +
  labs(
    title = "Standardized Morbidity Ratios (SMR)",
    subtitle = "SMR > 1 means higher risk than standard population",
    x = NULL,
    y = "SMR (Observed / Expected)"
  ) +
  theme(legend.position = "none")
```

::: {.fragment}
**Interpretation:** Men have `r sprintf("%.0f%%", (indirect_std$SMR[indirect_std$sex == "Man"] - 1) * 100)` higher risk than expected from standard rates
:::

## Direct Standardization: The Fix

**Goal:** What would rates be if men and women had the **same** age distribution?

```{r}
#| label: direct-std-sim2
#| echo: true

# Step 1: Use marginal P(Age) as standard weights
std_weights <- sim2_data %>%
  count(age_group) %>%
  mutate(weight = n / sum(n))

# Step 2: Apply to sex-age specific rates
direct_std <- age_sex_rates %>%
  left_join(std_weights, by = "age_group") %>%
  group_by(sex) %>%
  summarise(standardized_rate = sum(rate * weight))

direct_std
```

## Crude vs Standardized Rates

```{r}
#| label: crude-vs-std-sim2
#| echo: false

comparison_sim2 <- bind_rows(
  crude_sex %>% select(sex, rate = crude_rate) %>% mutate(type = "Crude"),
  direct_std %>% select(sex, rate = standardized_rate) %>% mutate(type = "Standardized")
)

ggplot(comparison_sim2, aes(x = sex, y = rate, fill = type)) +
  geom_col(position = "dodge", width = 0.7) +
  geom_text(aes(label = percent(rate, accuracy = 0.1)),
            position = position_dodge(width = 0.7),
            vjust = -0.5, size = 5) +
  scale_y_continuous(labels = percent_format(), limits = c(0, 0.30)) +
  scale_fill_manual(values = c("#E74C3C", "#27AE60")) +
  labs(
    title = "Crude vs Direct Standardized Rates",
    subtitle = "Standardization reveals men have higher rates",
    x = NULL,
    y = "Late Presentation Rate",
    fill = NULL
  ) +
  theme(legend.position = "top")
```

::: {.fragment}
**Pattern reverses!** Standardization reveals the truth: men have higher risk.
:::

## Key Insight: Confounding

::: {.incremental}
- **Crude rates get it wrong** because Sex → Age and Age → Outcome
- **Age is a confounder** of the Sex-Outcome relationship
- **Direct standardization** fixes $P(\text{Age})$ to break the confounding
- **Result:** Correct comparison of sex differences
:::

::: {.fragment}
<br>

**Next:** What happens when we have multiple confounders?
:::

# Example 3: Age + Sex + GOC {.section-break background-color="#2C3E50"}

## Adding Gender Opportunity Context: Multiple Confounders

**Research question:** Same question, but now we also have GOC data

GOC = Gender Opportunity Context (gendered norms/power/access)

::: {.incremental}
- **Known truth** (from simulation):
  - Age effect: Young=5%, Middle=15%, Older=40%
  - Sex effect: Men higher within age groups (constant logit-scale effect)
  - GOC effect: Low/Mid higher than High (constant logit-scale effects)
:::

---

## Multiple Confounding Pathways

::: {.incremental}
- GOC → Age (low GOC = younger; more constraints/shorter healthy lifespan)
- GOC → Outcome (low GOC = higher risk; barriers to access/norms)
- Sex → Age (women live longer)
- Age → Outcome
- Sex → Outcome
:::

::: {.fragment}
<br>

**The complexity:** All three variables affect both each other AND the outcome
:::

## Causal Structure: Multiple Confounders

```{r}
#| label: dag3
#| echo: false
#| fig-width: 10
#| fig-height: 5

dag3 <- dagify(
  Outcome ~ Age + Sex + GOC,
  Age ~ Sex + GOC,
  coords = list(
    x = c(GOC = 1, Sex = 1, Age = 2, Outcome = 3),
    y = c(GOC = 1, Sex = 3, Age = 2, Outcome = 2)
  )
)

ggdag(dag3, text_size = 5, node_size = 20) +
  theme_dag() +
  labs(title = "Gender Opportunity Context introduces multiple pathways")
```

::: {.fragment}
**This is complex!**

To compare groups fairly, we need to adjust for Age AND GOC.

Manual standardization becomes unmanageable.
:::

## Simulated Dataset 3: Age by Sex and GOC

```{r}
#| label: age-by-sex-goc-sim3
#| echo: false
#| fig-width: 12

sim3_data %>%
  count(sex, goc, age_group) %>%
  group_by(sex, goc) %>%
  mutate(prop = n / sum(n)) %>%
  ggplot(aes(x = age_group, y = prop, fill = age_group)) +
  geom_col() +
  geom_text(aes(label = percent(prop, accuracy = 1)),
            position = position_stack(vjust = 0.5), size = 3) +
  facet_grid(sex ~ goc) +
  scale_y_continuous(labels = percent_format()) +
  scale_fill_manual(values = c("#3498DB", "#E67E22", "#E74C3C")) +
  labs(
    title = "Age Distribution by Sex and GOC (Dataset 3)",
    subtitle = "Complex patterns: P(Age | Sex, GOC) varies, but marginal (i.e., overall) P(Age) same as D1/D2",
    x = "Age Group",
    y = "Proportion",
    fill = "Age"
  ) +
  theme(legend.position = "bottom")
```

## Crude Rates: Confounded

```{r}
#| label: crude-all-sim3
#| echo: true

# By Sex
sim3_data %>%
  group_by(sex) %>%
  summarise(crude_rate = mean(late_presentation))

# By GOC
sim3_data %>%
  group_by(goc) %>%
  summarise(crude_rate = mean(late_presentation))
```

::: {.fragment}
**All confounded!**
Low GOC appears to have lower risk (wrong!) because low GOC contexts are younger.
:::

## Why Manual Standardization Fails

To standardize by Age and GOC:

::: {.incremental}
- Need age-GOC-specific rates for each sex: 3 ages × 3 GOC = **9 strata per sex**
- With small cell counts, rates become unstable
- Adding more confounders (ethnicity, region, comorbidity) → exponential growth
- **Becomes unmanageable quickly**
:::

::: {.fragment}
<br>

**Solution:** Regression models do multivariable standardization automatically
:::

## Poisson Regression

```{r}
#| label: glm-sim3
#| echo: true

glm_fit <- glm(late_presentation ~ sex + age_group + goc,
               family = poisson(link = "log"),
               data = sim3_data)
```

::: {.incremental}
- Fits a model: $\log(E[Y \mid \text{Sex}, \text{Age}, \text{GOC}]) = \beta_0 + \beta_{\text{Sex}} + \beta_{\text{Age}} + \beta_{\text{GOC}}$
- Automatically adjusts for all confounders simultaneously
- **This is standardization!** Model weights by the data's covariate distribution
:::

---

## Adjusted Probabilities: Men vs Women

```{r}
#| label: emmeans-sim3
#| echo: true

library(emmeans)

# Get marginal probabilities by sex (adjusted for age and GOC)
emmeans(glm_fit, ~ sex, type = "response")

# Probability difference
pairs(regrid(emmeans(glm_fit, ~ sex, type = "response")), reverse = TRUE)
```

::: {.fragment}
**Men have 3.2 percentage points higher risk than women** (after adjusting for age and GOC)
:::

---

## Adjusted Probabilities by GOC

```{r}
#| label: emmeans-goc-sim3
#| echo: true

# Get marginal probabilities by GOC (adjusted for age and sex)
emmeans(glm_fit, ~ goc, type = "response")
```

---

## GOC Probability Differences

```{r}
#| label: emmeans-goc-contrasts
#| echo: true

# Probability differences
pairs(regrid(emmeans(glm_fit, ~ goc, type = "response")), reverse = TRUE)
```

::: {.fragment}
**Low GOC contexts have 4.7 pp higher risk than High GOC** (adjusted for age/sex)

**Mid GOC contexts have 1.7 pp higher risk than High GOC** (adjusted for age/sex)
:::

---

## What the Model Does

::: {.incremental}
- Model fits on log scale: $\log(E[Y \mid \text{Sex}, \text{Age}, \text{GOC}]) = \beta_0 + \beta_{\text{Sex}} + \beta_{\text{Age}} + \beta_{\text{GOC}}$
- `emmeans` converts those to **predicted probabilities** we can interpret
- Then calculates **probability differences** between groups
- Equivalent to standardizing by Age and GOC simultaneously
- **Much more manageable** than 9 strata of manual standardization
:::

# Summary {.section-break background-color="#2C3E50"}

## The Story

1. **Example 1 (Age only):** Crude rate depends on population structure
2. **Example 2 (Age + Sex):** Confounding creates Simpson's Paradox; direct standardization fixes it
3. **Example 3 (Age + Sex + GOC):** Multiple confounders: regression models generalize standardization

## All Methods Compared

```{r}
#| label: summary-table
#| echo: false

summary_comparison <- tribble(
  ~Method, ~`Example 2 (Sex)`, ~`Example 3 (GOC + Sex)`,

  "Crude",
  "Women appear worse",
  "Low GOC appears better (wrong!)",

  "Direct Standardization",
  "Men worse (correct)",
  "Too many strata (unmanageable)",

  "Poisson GLM",
  "Men worse (RR ≈ 1.06)",
  "Men worse, Low GOC worse (correct)"
)

summary_comparison %>%
  kable(caption = "Getting the Expected Value Right Requires Understanding Causal Structure")
```

## Key Takeaways

::: {.incremental}
1. **Expectation** = weighted average of subgroup rates
2. **Population structure** determines which weights we use
3. **Confounding** occurs when groups have different structures
4. **Standardization** uses common weights to make fair comparisons
5. **Regression models** = automated multivariable standardization
:::

## Connecting to Policy

If we **only** looked at crude rates:

::: {.incremental}
- **Example 2:** Crude data: women seem worse. Truth: men higher risk  
- **Example 3:** Crude data: low-opportunity contexts seem safer. Truth: risk highest where gender equality is lowest
- **Wrong interventions:** Misallocated resources, ineffective programs
- **Root cause:** Ignored causal structure and population composition
:::

::: {.fragment}

**Gendered health analysis requires understanding confounding, causal structure, and standardization**

:::

## Summary

**What you just learned:**

- Causal structure and confounding
- Direct/indirect standardization for 1-2 confounders
- Why manual standardization fails with many confounders

# Questions? {.center background-color="#2C3E50"}

Email me at `ewestlund@jhu.edu` and I'm happy to talk!

## Thank You

<br><br>

**Erik Westlund**

ewestlund@jhu.edu<br>
Johns Hopkins Biostatistics Center<br>
Johns Hopkins Bloomberg School of Public Health<br>

<br>

Slides and code: [github.com/JHBiostatCenter/geu-standardization-lecture](https://github.com/JHBiostatCenter/geu-standardization-lecture)
